Guaranteeing Safety and Liveness of Unmanned Aerial Vehicle Platoons on Air Highways

### 2016-05-10 docs/AIAA 2016 highway platooning/conclusions.tex
Made minor to medium changes according to Claire and Jaime's comments

% !TEX root = main.tex
\section{Conclusions}
To address the important and urgent problem of the traffic management of UAVs, we proposed to have platoons of UAVs traveling on air highways. We showed how such an airspace structure leads to a much easier safety and liveless analysis. We provided simulations which show that by putting vehicles into platoons, many complex maneuvers can be performed using just a few different reachable sets.

>>>>

% !TEX root = main.tex
\section{Conclusions}
To address the important and urgent problem of the traffic management of UAVs, we proposed to have platoons of UAVs traveling on air highways. We showed how such an airspace structure leads to a much easier safety and liveness analysis. We provided simulations which show that by organizing vehicles into platoons, many complex maneuvers can be performed using just a few different reachable sets.

### 2016-05-10 docs/AIAA 2016 highway platooning/high_eikon.tex
Made minor to medium changes according to Claire and Jaime's comments

The Eikonal equation \eqref{eq:eikonal} can be efficiently computed numerically using the fast marching method \cite{Sethian96}.

>>>>

The Eikonal equation \eqref{eq:eikonal} can be efficiently computed numerically using the fast marching method \cite{Sethian96}; each computation takes on the order of merely a second.

### 2016-05-10 docs/AIAA 2016 highway platooning/high_form.tex
Made minor to medium changes according to Claire and Jaime's comments

\begin{equation}
\label{eq:ahpp} % air highway placement problem
\begin{aligned}
& \min_{\hws_N, N} \sum_{i=1}^N \int_0^1 \cmap(\hw_i(s)) ds \\
& \text{subject to \eqref{eq:hw_seq}} 
\end{aligned}
\end{equation}

In other words, we consider air highways to be line segments of constant altitude over a region, and UAV platoons travel on these air highways to get from some origin to some destination. Any UAV flying on a highway over some position $\pos$ incurs a cost of $\cmap(\pos)$, so that the total cost of flying from the origin to the destination is given by \eqref{eq:ahpp}. The air highway placement problem minimizes the cumulative cost of flying from some origin $p^o$ to some destination $p^d$ along the sequence of highways $\hws_N=\{\hw_1, \hw_2, \ldots, \hw_N\}$.

>>>>

\begin{equation}
\label{eq:ahpp} % air highway placement problem
\begin{aligned}
& \min_{\hws_N, N} \left\{\left(\sum_{i=1}^N \int_0^1 \cmap(\hw_i(s)) ds\right) + R(N)\right\}\\
& \text{subject to \eqref{eq:hw_seq}} 
\end{aligned}
\end{equation}

\noindent where $R(\cdot)$ is a regularizer such as $R(N) = N^2$.

In other words, we consider air highways to be line segments of constant altitude over a region, and UAV platoons travel on these air highways to get from some origin to some destination. Any UAV flying on a highway over some position $\pos$ incurs a cost of $\cmap(\pos)$, so that the total cost of flying from the origin to the destination is given by \eqref{eq:ahpp}. The air highway placement problem minimizes the cumulative cost of flying from some origin $p^o$ to some destination $p^d$ along the sequence of highways $\hws_N=\{\hw_1, \hw_2, \ldots, \hw_N\}$. The regularization term is used to prevent $N$ from being arbitrarily large.

### 2016-05-10 docs/AIAA 2016 highway platooning/high_sim.tex
Made the "easy" changes to AIAA paper according to Claire's comments
Made minor to medium changes according to Claire and Jaime's comments

This assumption assigns costs in descending order to the categories ``regions around airports'', ``cities'', ``other'', and ``water''. Flying a UAV in each category is more costly by a factor of $b$ compared to the next most important category. The factor $b$ is a tuning parameter that we adjusted to vary the relative importance of the different categories, and we used $b=4$ in the figures below.

Figure \ref{fig:airHighway_results} shows the San Francisco Bay Area (geographic) map, cost map, cost-minimizing paths, and contours of the value function. The region enclosed by the black boundary represent ``region around airports'', which have the highest cost. The dark blue, yellow, and light blue regions represent the ``cities'', the ``water'', and the ``other''' categories, respectively. We assumed that the origin corresponds to the city ``Concord'', and chose a number of other major cities as destinations.

A couple of important observations can be made here. First, the cost-minimizing paths to the various destinations in general overlap, and only split up when they are very close to entering their destination cities. This is intuitively desirable because the number of air highways is kept low. In addition, the cost of flying in the airspace according to the cost map is minimized. Secondly, the spacing of the contours, which correspond to level curves of the value function, have a spacing corresponding to the cost map. This provides insight into the placement of air highways to destinations that were not shown in this example.

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{"fig/airHighway_results"}
	\caption{Cost-minimizing paths computed by the Fast Marching Method based on the assumed cost map of the San Francisco Bay Area}
	\label{fig:airHighway_results}
\end{figure}

Fig. \ref{fig:airHighway_sparse} shows the result of converting the cost-minimizing paths to a small number of waypoints. The left plot shows the waypoints, interpreted as the start and end points of air highways, over a white background for clarity. The right plot shows these air highways over the map of the Bay Area.

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{"fig/airHighway_sparse"}
	\caption{Results of conversion from cost-minimizing paths to highway way points.}
	\label{fig:airHighway_sparse}
\end{figure}

>>>>

This assumption assigns costs in descending order to the categories ``regions around airports'', ``cities'', ``other'', and ``water''. Flying a UAV in each category is more costly by a factor of $b\ge1$ compared to the next most important category. The factor $b$ is a tuning parameter that we adjusted to vary the relative importance of the different categories, and we used $b=4$ in the figures below.

Figure \ref{fig:airHighway_results} shows the San Francisco Bay Area (geographic) map, cost map, cost-minimizing paths, and contours of the value function. The region enclosed by the black boundary represents ``region around airports'', which have the highest cost. The dark blue, yellow, and light blue regions represent the ``cities'', the ``water'', and the ``other''' categories, respectively. We assumed that the origin corresponds to the city ``Concord'', and chose a number of other major cities as destinations.

A couple of important observations can be made here. First, the cost-minimizing paths to the various destinations in general overlap, and only split up when they are very close to entering their destination cities. This is intuitively desirable because the number of air highways is kept low. In addition, the cost of flying in the airspace according to the cost map is minimized. Secondly, the contours, which correspond to level curves of the value function, have a spacing corresponding to the cost map. This provides insight into the placement of air highways to destinations that were not shown in this example.

\begin{figure}
	\centering
	\includegraphics[width=0.95\textwidth]{"fig/airHighway_results"}
	\caption{Cost-minimizing paths computed by the Fast Marching Method based on the assumed cost map of the San Francisco Bay Area}
	\label{fig:airHighway_results}
\end{figure}

Figure \ref{fig:airHighway_sparse} shows the result of converting the cost-minimizing paths to a small number of waypoints. The left plot shows the waypoints, interpreted as the start and end points of air highways, over a white background for clarity. The right plot shows these air highways over the map of the Bay Area.

\begin{figure}
	\centering
	\includegraphics[width=0.95\textwidth]{"fig/airHighway_sparse"}
	\caption{Results of conversion from cost-minimizing paths to highway way points.}
	\label{fig:airHighway_sparse}
\end{figure}

\subsection{Real-Time Highway Location Updates}
Since \eqref{eq:eikonal} can be solved in approximately a second, the air highway placement process can be redone in real-time if the cost map changes. This flexibility can be useful in any situation in which unforeseen circumstances cause a change in the cost of a particular region of the airspace. For example, accidents or disaster response vehicles may result in an area temporarily having a high cost. On the other hand, depending on for instance the time of day, it may be most desirable to fly in different regions of the airspace, resulting in those regions temporarily having a low cost.

### 2016-05-10 docs/AIAA 2016 highway platooning/high_waypoints.tex
Made minor to medium changes according to Claire and Jaime's comments

% !TEX root = main.tex
\subsection{From Paths to Waypoints}
Each of the cost-minimizing paths $\ppath_i$ computed from the solution to the Eikonal equation consists of a continuous set of points. Each path $\ppath_i$ is an approximation to the sequence of highways $\hws_{N_i}^i = \{\hw^i_j\}_{i=1,j=1}^{i=M,j=N_i}$ defined in \eqref{eq:ahpp}, but now indexed by the corresponding path index $i$.

@@

If there is a large change in heading within a small section of the cost-minimizing path, then the collection of point may contain many points which are close together. In addition, there may be multiple paths that are very close to each other (in fact, this behavior is desirable), which may contribute to cluttering the airspace with too many waypoints. We propose to sparsify the waypoints by clustering the points. After clustering, we replace each cluster of points with a single point located at the centroid of the cluster.

>>>>

% !TEX root = main.tex
\subsection{From Paths to Waypoints}
Each of the cost-minimizing paths $\ppath_i$ computed from the solution to the Eikonal equation consists of a closely-spaced set of points. Each path $\ppath_i$ is an approximation to the sequence of highways $\hws_{N_i}^i = \{\hw^i_j\}_{i=1,j=1}^{i=M,j=N_i}$ defined in \eqref{eq:ahpp}, but now indexed by the corresponding path index $i$.

@@

If there is a large change in heading within a small section of the cost-minimizing path, then the collection of points may contain many points which are close together. In addition, there may be multiple paths that are very close to each other (in fact, this behavior is desirable), which may contribute to cluttering the airspace with too many waypoints. We propose to sparsify the waypoints by clustering the points. After clustering, we replace each cluster of points with a single point located at the centroid of the cluster.

### 2016-05-10 docs/AIAA 2016 highway platooning/introduction.tex
Made minor to medium changes according to Claire and Jaime's comments

% !TEX root = main.tex
\section{Introduction}
Unmanned aerial vehicle (UAV) systems have in the past been mainly used for military operations \cite{Tice91}. Recently, however, there has been an immense surge of interest in using UAVs for civil applications. Through projects such as Amazon Prime Air \cite{PrimeAir} and Google Project Wing \cite{ProjectWing}, companies are looking to send UAVs into the airspace to not only deliver commerical packages, but also for important tasks such as aerial surveillance, emergency supply delivery, and search and rescue. In the future, the applications of UAVs are only limited by human imagination.

>>>>

% !TEX root = main.tex
\section{Introduction}
\MCnote{Add reference suggested by Claire}
\MCnote{Be consistent with UAS vs UAV}
Unmanned aerial systems (UAS) have in the past been mainly used for military operations \cite{Tice91}. Recently, however, there has been an immense surge of interest in using UASs for civil applications. Through projects such as Amazon Prime Air \cite{PrimeAir} and Google Project Wing \cite{ProjectWing}, companies are looking to send UAVs into the airspace to not only deliver commercial packages, but also for important tasks such as aerial surveillance, emergency supply delivery, and search and rescue. In the future, the applications of UAVs are only limited by human imagination.

### 2016-05-10 docs/AIAA 2016 highway platooning/plat_controllers.tex
Made the "easy" changes to AIAA paper according to Claire's comments
Made minor to medium changes according to Claire and Jaime's comments

% !TEX root = main.tex
\subsection{Reachability-Based Controllers \label{sec:reach_ctrl}}
Reachability analysis is useful for constructing controllers in a large variety of situations. In order to construct different controllers, an appropriate target set needs to be defined depending on the goal of the controller. If one defines the target set to be a set of desired states, the reachable set, once computed, would represent the states from which a system needs to first arrive at in order to reach the desired states. On the other hand, if the target set represents a set of undesirable states, then the reachable set would indicate the region of the state space that the system needs to avoid. In addition, the system dynamics with which the reachable set is computed provides additional flexibility when using reachability to construct controllers.

@@

\subsubsection{Getting to a Target State \label{sec:abs_target_ctrl}}
The controller used by a vehicle to reach a target state is important in a couple of situations in the platooning context. First, a vehicle in the ``Free'' mode can use the controller to merge onto a highway, forming a platoon and changing modes to a ``Leader'' vehicle. Second, a vehicle in either the ``Leader'' mode or the ``Follower'' mode can use this controller to change to a different highway, changing modes to a ``Leader'' vehicle.

In both of the above cases, we use the dynamics of a single vehicle specified in \eqref{eq:dyn}. The target state would be a position $(\Pos_x, \Pos_y)$ representing the desired merging point on the highway, along with a velocity $(\Vel_x, \Vel_y)$ that corresponds to the speed and direction of travel specified by the highway. For the reachability computation, we define the target set to be a small range of states around the target state $x_H = (\Pos_x, \Pos_y, \Vel_x, \Vel_y)$:

@@

\begin{enumerate}
\item Move towards $\bar{x}_H$ in a straight line, with some velocity, until $V_H(-T,x)\le 0$. In practice, this step consistently drives the system into the reachable set.
\item Apply the optimal control extracted from $V_H(-T,x)$ according to \eqref{eq:HJB_ctrl_syn} until $\mathcal{L}_H$ is reached.
\end{enumerate}

@@

We can now define the implicit surface function $l_S(x)$ corresponding to $\mathcal{L}_S$, and solve the HJI PDE \eqref{eq:HJIPDE} using $l_S(x)$ as the terminal condition. As before, the zero sublevel set of the solution $V_S(t,x)$ specifies the reachable set $\mathcal{V}_S(t)$, which characterizes the states in the augmented relative coordinates, as defined in \eqref{eq:rel_dyn_aug}, from which $Q_i$ \textit{cannot} avoid $\mathcal{L}_S$ for a time period of $t$, if $Q_j$ uses the worst case control. To avoid collisions, $Q_i$ must apply the safety controller according to \eqref{eq:HJI_ctrl_syn} on the boundary of the reachable set in order to avoid going into the reachable set. The following algorithm wraps our safety controller around liveness controllers:

\begin{enumerate}
\item For a specified time horizon $t$, evaluate$V_S(-t,x_i-x_j)$ for all $j\in \mathcal{Q}(i)$.

$\mathcal{Q}(i)$ is the set of quadrotors with which quadrotor $i$ checks safety against.
\item Use the safety or liveness controller depending on the values $V_S(-t,x_i-x_j),j\in \mathcal{Q}(i)$:

@@

\subsubsection{Following a Platoon} \label{sec:follow_platoon}
Follower vehicles use a feedback control law tracking a nominal position and velocity in the platoon, with an additional feed-forward term given by the leader's acceleration input; here, for simplicity, we assume perfect communication between the leader and the follower vehicles. This following law enables smooth vehicle trajectories in the relative platoon frame, while allowing the platoon as a whole to perform agile maneuvers by transmitting the leader's acceleration command $u_{P_1}(t)$ to all vehicles.

The $i$-th member of the platoon, $Q_{P_i}$, is expected to track a relative position in the platoon $r^i = (r_x^i,r_y^i)$ with respect to the leader's position $p_{P_1}$, and the leader's velocity $v_{P_1}$ at all times. The resulting control law has the form:
\begin{equation}\label{eq:follow}
u^i(t) = k_p \big[p_{P_1}(t) + r^i(t) - p^i(t) \big] + k_v\big[v_{P_1}(t) - v^i(t)\big] + u_{P_1}(t)
\end{equation}
for some $k_p,k_v>0$. In particular, a simple rule for determining $r^i(t)$ in a single-file platoon is given for $Q_{P_i}$ as:
\begin{equation}\label{eq:nominal_pos}
r^i(t) = - (i-1) b \frac{v_{P_1}}{\|v_{P_1}\|_2}
\end{equation}
where $b$ is the spacing between vehicles along the platoon and $\frac{v_{P_1}}{\|v_{P_1}\|_2}$ is the platoon leader's direction of travel.

>>>>

% !TEX root = main.tex
\subsection{Reachability-Based Controllers \label{sec:reach_ctrl}}
Reachability analysis is useful for constructing controllers in a large variety of situations. In order to construct different controllers, an appropriate target set needs to be defined depending on the goal of the controller. If one defines the target set to be a set of desired states, the reachable set would represent the states that a system needs to first arrive at in order to reach the desired states. On the other hand, if the target set represents a set of undesirable states, then the reachable set would indicate the region of the state space that the system needs to avoid. In addition, the system dynamics with which the reachable set is computed provide additional flexibility when using reachability to construct controllers.

@@

\subsubsection{Getting to a Target State \label{sec:abs_target_ctrl}}
The controller used by a vehicle to reach a target state is important in two situations in the platooning context. First, a vehicle in the ``Free'' mode can use the controller to merge onto a highway, forming a platoon and changing modes to a ``Leader'' vehicle. Second, a vehicle in either the ``Leader'' mode or the ``Follower'' mode can use this controller to change to a different highway, becoming a ``Leader'' vehicle.

In both of the above cases, we use the dynamics of a single vehicle specified in \eqref{eq:dyn}. The target state would be a position $(\Pos_x, \Pos_y)$ representing the desired merging point on the highway, along with a velocity $(\Vel_x, \Vel_y)$ that corresponds to the speed and direction of travel specified by the highway. For the reachability computation, we define the target set to be a small range of states around the target state $\bar x_H = (\Pos_x, \Pos_y, \Vel_x, \Vel_y)$:

@@

\begin{enumerate}
\item Move towards $\bar{x}_H$ in pure pursuit with some velocity, until $V_H(-T,x)\le 0$. In practice, this step consistently drives the system into the reachable set.
\item Apply the optimal control extracted from $V_H(-T,x)$ according to \eqref{eq:HJB_ctrl_syn} until $\mathcal{L}_H$ is reached.
\end{enumerate}

@@

We can now define the implicit surface function $l_S(x)$ corresponding to $\mathcal{L}_S$, and solve the HJI PDE \eqref{eq:HJIPDE} using $l_S(x)$ as the terminal condition. As before, the zero sublevel set of the solution $V_S(t,x)$ specifies the reachable set $\mathcal{V}_S(t)$, which characterizes the states in the augmented relative coordinates, as defined in \eqref{eq:rel_dyn_aug}, from which $Q_i$ \textit{cannot} avoid $\mathcal{L}_S$ for a time period of $t$, if $Q_j$ uses the worst-case control. To avoid collisions, $Q_i$ must apply the safety controller according to \eqref{eq:HJI_ctrl_syn} on the boundary of the reachable set in order to avoid going into the reachable set. The following algorithm wraps our safety controller around liveness controllers:

\begin{enumerate}
\item For a specified time horizon $t$, evaluate $V_S(-t,x_i-x_j)$ for all $j\in \mathcal{Q}(i)$.

$\mathcal{Q}(i)$ is the set of quadrotors with which quadrotor $i$ checks safety.
\item Use the safety or liveness controller depending on the values $V_S(-t,x_i-x_j),j\in \mathcal{Q}(i)$:

@@

\subsubsection{Following a Platoon} \label{sec:follow_platoon}
Follower vehicles use a feedback control law tracking a nominal position and velocity in the platoon, with an additional feedforward term given by the leader's acceleration input; here, for simplicity, we assume perfect communication between the leader and the follower vehicles. This following law enables smooth vehicle trajectories in the relative platoon frame, while allowing the platoon as a whole to perform agile maneuvers by transmitting the leader's acceleration command $u_{P_1}(t)$ to all vehicles.

The $i$-th member of the platoon, $Q_{P_i}$, is expected to track a relative position in the platoon $r^i = (r_x^i,r_y^i)$ with respect to the leader's position $p_{P_1}$, and the leader's velocity $v_{P_1}$ at all times. The resulting control law has the form:
\begin{equation}\label{eq:follow}
u^i(t) = k_p \big[p_{P_1}(t) + r^i(t) - p^i(t) \big] + k_v\big[v_{P_1}(t) - v^i(t)\big] + u_{P_1}(t)
\end{equation}
for some $k_p,k_v>0$. In particular, a simple rule for determining $r^i(t)$ in a single-file platoon is given for $Q_{P_i}$ as:
\begin{equation}\label{eq:nominal_pos}
r^i(t) = - (i-1) \sepdist \frac{v_{P_1}}{\|v_{P_1}\|_2}
\end{equation}
where $b$ is the spacing between vehicles along the platoon and $\frac{v_{P_1}}{\|v_{P_1}\|_2}$ is the platoon leader's direction of travel.

### 2016-05-10 docs/AIAA 2016 highway platooning/plat_form.tex
Made the "easy" changes to AIAA paper according to Claire's comments
Made minor to medium changes according to Claire and Jaime's comments

% !TEX root = main.tex
\section{Unmanned Aerial Vehicle Platooning \label{sec:platooning}}
\subsection{UAVs in Platoons}
\subsubsection{Vehicle Dynamics}
Consider a UAV whose dynamics are given by
\begin{equation}
\dot{x} = f(x,u)
\end{equation}

@@

\item Leader: 
A Leader vehicle is the vehicle at the front of a platoon (which could consist of only the vehicle itself). The available maneuvers and mode transitions are

@@

If a vehicle from any of the other modes malfunctions, it transitions into the Faulty mode and descends after some pre-specified duration $t_\text{faulty}$.
\end{itemize}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{"fig/vehicleModes"}
	\caption{Hybrid modes for vehicles in platoons.}
	\label{fig:vehicleModes}
\end{figure}

Suppose that there are $N$ vehicles in total. We will denote the $N$ vehicles as $\veh{i}, i=1\ldots,N$. We consider a platoon of vehicles to be a group of $M\le N$ vehicles, denoted $\veh{P_1}, \ldots, \veh{P_M}, \{P_j\}_{j=1}^M \subseteq \{i\}_{i=1}^N$, in a single-file formation; for convenience, we also denote $P_0 = P_{M+1} = \emptyset$. When necessary, we will use superscripts to denote vehicles of different platoons: $\veh{P_i^j}$ represents the $i$th vehicle in the $j$th platoon.

For convenience, let $\vehSCS{i}$ denote the set of indices of vehicles with respect to which $Q_i$ checks safety against. If vehicle $\veh{i}$ is a free vehicle, then it must check for safety with respect to all other vehicles, $\vehSCS{i} = \{j: j\neq i\}$. If the vehicle is part of a platoon, then it checks safety with respect to the platoon member in front and behind, $\vehSCS{i} = \{P_{j+1}, P_{j-1}\}$. Figure \ref{fig:vehicleNotation} summarizes the indexing system of the vehicles.

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{"fig/vehicleNotation"}
	\caption{Notation for vehicles in platoons.}
	\label{fig:vehicleNotation}
\end{figure}

We will assume that the vehicles in a platoon travel along an air highway. The vehicles maintain a separation distance of $b$ with its neighbors inside the platoon. In order to allow for close proximity of the vehicles and the ability to resolve multiple simultaneous safety breaches, we assume that in the event of a malfunction, a vehicle will be able to exit the altitude range of the highway within a duration of $t_\text{faulty}=1.5$. Such a requirement may be implemented practically as an emergency landing procedure to which the vehicles revert when a malfunction is detected.

@@

\begin{itemize}
\item merge onto a highway to form a platoon
\item join a new or different platoon
\item leave a platoon to create a new one
\item reacting to malfunctioning or intruder vehicles
\end{itemize}

We also propose more basic controllers to perform other simpler actions such as
\begin{itemize}
\item following the highway at constant altitude at a specified speed
\item maintaining a constant relative position and velocity with the leader of a platoon
\end{itemize}

In general, the control strategies of each vehicle have a liveness component, which specifies a set of states towards which the vehicle aims to reach, and a safety component, which specifies a set of states that it must avoid. Together, the liveness and safety controllers guarantee the success and safety of a vehicle in the airspace making any desired mode transition. In this paper, these guarantees are provided using reachability analysis, and allow the multi-UAV system to perform joint maneuvers essential to maintaining structure in the airspace.

>>>>

% !TEX root = main.tex
\section{Unmanned Aerial Vehicle Platooning \label{sec:platooning}}
\MCnote{Add introduction section}
\subsection{UAVs in Platoons}
\subsubsection{Vehicle Dynamics}
Consider a UAV whose dynamics are given by
\begin{equation}
\dot{x} = f(x,u)
\end{equation}

@@

\item Leader:

A Leader vehicle is the vehicle at the front of a platoon (which could consist of only the vehicle itself). The available maneuvers and mode transitions are

@@

If a vehicle from any of the other modes malfunctions, it transitions into the Faulty mode. A Faulty vehicle is assumed to descend via a fail-safe mechanism after some pre-specified duration $t_\text{faulty}$ to a different altitude level where it no longer poses a threat to vehicles on the air highway system.
\end{itemize}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.95\textwidth]{"fig/vehicleModes"}
	\caption{Hybrid modes for vehicles in platoons.}
	\label{fig:vehicleModes}
\end{figure}

Suppose that there are $N$ vehicles in total in the airspace containing the highway system. We will denote the $N$ vehicles as $\veh{i}, i=1\ldots,N$. We consider a platoon of vehicles to be a group of $M\le N$ vehicles, denoted $\veh{P_1}, \ldots, \veh{P_M}, \{P_j\}_{j=1}^M \subseteq \{i\}_{i=1}^N$, in a single-file formation; for convenience, we also denote $P_0 = P_{M+1} = \emptyset$. When necessary, we will use superscripts to denote vehicles of different platoons: $\veh{P_i^j}$ represents the $i$th vehicle in the $j$th platoon.

For convenience, let $\vehSCS{i}$ denote the set of indices of vehicles with respect to which $Q_i$ checks safety. If vehicle $\veh{i}$ is a free vehicle, then it must check for safety with respect to all other vehicles, $\vehSCS{i} = \{j: j\neq i\}$. If the vehicle is part of a platoon, then it checks safety with respect to the platoon member in front and behind, $\vehSCS{i} = \{P_{j+1}, P_{j-1}\}$. Figure \ref{fig:vehicleNotation} summarizes the indexing system of the vehicles.

\begin{figure}
	\centering
	\includegraphics[width=0.95\textwidth]{"fig/vehicleNotation"}
	\caption{Notation for vehicles in platoons.}
	\label{fig:vehicleNotation}
\end{figure}

We will organize the vehicles into platoons travel along air highways. The vehicles maintain a separation distance of $\sepdist$ with their neighbors inside the platoon. In order to allow for close proximity of the vehicles and the ability to resolve multiple simultaneous safety breaches, we assume that in the event of a malfunction, a vehicle will be able to exit the altitude range of the highway within a duration of $t_\text{faulty}=1.5$. Such a requirement may be implemented practically as an fail-safe mechanism to which the vehicles revert when a malfunction is detected.

@@

\begin{itemize}
\item merge onto a highway to form a platoon
\item join a new or different platoon
\item leave a platoon to create a new one
\item react to malfunctioning or intruder vehicles
\end{itemize}

We also propose more basic controllers to perform other simpler actions such as
\begin{itemize}
\item follow the highway at constant altitude at a specified speed
\item maintain a constant relative position and velocity with respect to the leader of a platoon
\end{itemize}

In general, the control strategies of each vehicle have a liveness component, which specifies a set of states that the vehicle aims to reach, and a safety component, which specifies a set of states that it must avoid. Together, the liveness and safety controllers guarantee the success and safety of a vehicle in the airspace making any desired mode transition. In this paper, these guarantees are provided using reachability analysis, and allow the multi-UAV system to perform joint maneuvers essential to maintaining structure in the airspace.

### 2016-05-10 docs/AIAA 2016 highway platooning/plat_reach.tex
Made the "easy" changes to AIAA paper according to Claire's comments
Made minor to medium changes according to Claire and Jaime's comments

\begin{equation}
\begin{aligned}
\mathcal{V}(t) &:= \{x\in\mathbb{R}^n \mid \exists \gamma\in\Gamma \text{ such that} \forall u_1(\cdot)\in\mathbb{U}_1, \\
&\exists s \in [t,0], \xi_f(s; t, x, u_1(\cdot), \gamma[u_1](\cdot)) \in \mathcal{L} \}
\end{aligned}
\end{equation}
where $\xi_f$ is the trajectory of the system satisfying initial conditions $\xi_f(t; x, t, u_1(\cdot), u_2(\cdot))=x$ and the following differential equation almost everywhere on $[-t, 0]$
\begin{equation}
\begin{aligned}
\frac{d}{ds}&\xi_f(s; x, t, u_1(\cdot), u_2(\cdot)) \\
&= f(\xi_f(s; x, t, u_1(\cdot), u_2(\cdot)), u_1(s), u_2(s))
\end{aligned}
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{"fig/RSExample"}
	\caption{Illustration of a target set, reachable set, and their implicit surface functions.}
	\label{fig:RSExample}
\end{figure}

@@

For our application, we will use several decoupled system models and utilize the decoupled HJ formulation in \cite{Chen15}, which enables real time 4D reachable set computations and tractable 6D reachable set computations.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.25\textwidth]{"fig/rel_coords"}
	\caption{Relative coordinates}
	\label{fig:rel_coords}
\end{figure}

>>>>

\begin{equation}
\begin{aligned}
\mathcal{V}(t) &:= \{x\in\mathbb{R}^n \mid \exists \gamma\in\Gamma \text{ such that } \forall u_1(\cdot)\in\mathbb{U}_1, \\
&\exists s \in [t,0], \xi_f(s; t, x, u_1(\cdot), \gamma[u_1](\cdot)) \in \mathcal{L} \}
\end{aligned}
\end{equation}
where $\xi_f$ is the trajectory of the system satisfying initial conditions $\xi_f(t; x, t, u_1(\cdot), u_2(\cdot))=x$ and the following differential equation almost everywhere on $[-t, 0]$
\begin{equation}
\begin{aligned}
\frac{d}{ds}&\xi_f(s; x, t, u_1(\cdot), u_2(\cdot)) \\
&= f(\xi_f(s; x, t, u_1(\cdot), u_2(\cdot)), u_1(s), u_2(s))
\end{aligned}
\end{equation}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.95\textwidth]{"fig/RSExample"}
	\caption{Illustration of a target set, reachable set, and their implicit surface functions.}
	\label{fig:RSExample}
\end{figure}

@@

For our application, we will use several decoupled system models and utilize the decoupled HJ formulation in \cite{Chen15}, which enables real-time 4D reachable set computations and tractable 6D reachable set computations.

@@

%\begin{figure}
%	\centering
%	\includegraphics[width=0.25\textwidth]{"fig/rel_coords"}
%	\caption{Relative coordinates}
%	\label{fig:rel_coords}
%\end{figure}

### 2016-05-10 docs/AIAA 2016 highway platooning/plat_safety.tex
Made minor to medium changes according to Claire and Jaime's comments

Within a duration of $\td$, there is a small chance that additional safety breaches may occur. However, as long as the total number of safety breaches does not exceed the number of affected quadrotors, collision avoidance of all the vehicles can be guaranteed for the duration $\td$. However, as our simulation results show, putting vehicles in single-file platoons makes the likelihood of multiple safety breaches low during a vehicle's malfunction and during the presence of one intruder vehicle.

>>>>

Within a duration of $\td$, there is a small chance that additional safety breaches may occur. However, as long as the total number of safety breaches does not exceed the number of affected quadrotors, collision avoidance of all the vehicles can be guaranteed for the duration $\td$. However, as our simulation results show, placing vehicles in single-file platoons makes the likelihood of multiple safety breaches low during a vehicle's malfunction and during the presence of one intruder vehicle.
